---
title: "data.data"
output:
  html_document:
    df_print: paged
---

# Libraries used

```{r Libraries}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2) # For plotting
library(dplyr) # For data handling
library(readr)
library(survival) # Main library, used for implementing Cox Modeling
library(pROC) # For plotting ROC curves
library(survminer)
library(utils) #For not having to upload the txt each time
library(riskRegression) # For performance indicators
```

# Dataset: Uploading and Exploratory Analysis

```{r Importing data from shared GitHub}

#setwd('~/Desktop/Data Science/Data Science Lab in Medicine/Big Data in Health Care/')
url <- "https://raw.githubusercontent.com/darthgween/BigDataInHealthCareProject/main/heart.valve.txt?token=GHSAT0AAAAAACR3GYSXO5Z3ARKVQY2S6SYSZRYZQWA"
data <- read.table(url, header = TRUE)
head(data)
```

```{r Dimensions}
dim(data)
```

Description of dataset variables:

-   ***Paz.id***: patient identification number.

-   ***log.lvmi***: natural logarithm of the “Left Ventricular Mass Index” (standardized) measured at baseline.

-   ***fuyrs***: follow-up time from surgery, in years.

-   ***status***: event indicator (1 = dead; 0 = censored).

-   ***sex***: sex of the patient (0 = M; 1 = F).

-   ***age***: age of the patient (years) at surgery.

-   ***con.cabg***: presence of coronary bypass (1 = yes; 0 = no).

-   ***creat***: pre-operative serum creatinine (µmol/mL).

-   ***lv***: pre-operative left ventricular ejection fraction (1 = high, 2 = moderate, 3 = low).

-   ***sten.reg.mix***: hemodynamics of the heart valve (1 = stenosis, 2 = regurgitation, 3 = mixed).

```{r Check for data structure and types}
str(data)
```

```{r Dataset Summary for all variables}
summary(data)
```

## Age distribution
[SOME COMMENT HERE]

```{r Age Histogram}
ggplot(data, aes(x = age)) +
  geom_histogram(binwidth = 5, fill = "lightblue", color = "black") +
  ggtitle("Distribuzione dell'età") +
  xlab("Età") +
  ylab("Frequenza")
```

## Correlation between variables

```{r Correlation Matrix}
cor_matrix <- cor(data, use = "complete.obs")
print(cor_matrix)

if (!require(corrplot)) {
    install.packages("corrplot")
    library(corrplot)
}
corrplot(cor_matrix, method = "circle")
```
```{r Transforming variables into Categoric}

data$sex <- factor(data$sex, levels = c(0, 1), labels = c("Male", "Female"))
data$con.cabg <- factor(data$con.cabg, levels = c(0, 1), labels = c("No", "Yes"))
data$lv <- factor(data$lv, levels = c(1, 2, 3), labels = c("High", "Moderate", "Low"))
data$sten.reg.mix <- factor(data$sten.reg.mix, levels = c(1, 2, 3), labels = c("Stenosis", "Regurgitation", "Mixed"))

```

## Deep Dive: Left Ventricular Mass Index

Il Left Ventricular Mass Index (LVMI) è un parametro clinico utilizzato per valutare la massa del ventricolo sinistro del cuore in relazione alla superficie corporea del paziente. È una misura importante nella diagnosi e nel monitoraggio di condizioni cardiache come l'ipertrofia ventricolare sinistra, che è un ispessimento delle pareti del ventricolo sinistro del cuore.

```{r Boxplot log.lvmi - Status}
ggplot(data, aes(x = factor(status), y = log.lvmi, fill = factor(status))) +
  geom_boxplot() +
  ggtitle("Boxplot of log.lvmi for Status") +
  xlab("Status") +
  ylab("log.lvmi") +
  scale_fill_discrete(name = "Status")

```


```{r Boxplot log.lvmi - Age}
ggplot(data, aes(x = age, y = log.lvmi)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  ggtitle("Scatter Plot: età vs log.lvmi") +
  xlab("Età") +
  ylab("log.lvmi")

```

# Models

## Univariate Models
Here we cycle through all variables of the dataset, excluding: the *patient ID* variable (which cannot possibly explain anything about the survival probability of patients), the *status* variable (which is the outcome) and the *fuyrs* variable (as it describes the follow-up time - hence, the censoring time).


```{r Univariate Function loop}
# The outcome variable is a survival object created with the time to event (fuyrs) and status indicator (status)
surv_obj <- Surv(data$fuyrs, data$status)

# Perform Cox regression for each predictor
# Excluding paz.id from the analysis as it's an identifier, not a predictor
variables <- c("log.lvmi", "sex", "age", "con.cabg", "creat", "lv", "sten.reg.mix")

# Store results in a list for easy access
univariate_models <- list()

for (var in variables) {
  # Formula to dynamically create the regression model
  formula <- as.formula(paste("surv_obj ~", var))
  
  # Fit the Cox model
  model <- coxph(formula, data = data)
  
  # Print the summary of the model
  print(summary(model))
  
  # Store the model in the list
  univariate_models[[var]] <- model
}
```
Belowa list of R code snippets is provided, in order to easily access each Univariate model, independently.

```{r Univariate Left Ventricular Mass Index}
univariate_models[["log.lvmi"]]
```

```{r Univariate Sex}
univariate_models[["sex"]]
```

```{r Univariate Age}
univariate_models[["age"]]
```

```{r Univariate Coronary Bypass}
univariate_models[["con.cabg"]]
```


```{r Univariate Creatinine}
univariate_models[["creat"]]
```


```{r Univariate Left Ventricular Ejection Fraction}
univariate_models[["lv"]]
```


```{r Univariate Hemodynamics of Heart Valve}
univariate_models[["sten.reg.mix"]]
```
## Base model

```{r Base Model}
base_model <- coxph(Surv(fuyrs, status) ~ sex + age + con.cabg + creat + lv + sten.reg.mix, data = data, x=T)

summary(base_model)
```

The Cox proportional hazards model shows significant effects of age, coronary artery bypass grafting (con.cabg), left ventricular function (lv), and stenosis severity (sten.reg.mix) on survival time. In particular:
1. Sex: the variable sex does not show a significant effect (p-value = 0.392228) on survival. The hazard ratio for females (1.317) would suggest a higher risk compared to males, but the wide confidence interval (0.7009 to 2.475) and high p-value indicate that this effect is not statistically robust. In many clinical studies, sex may influence outcomes due to differences in biological responses or treatment approaches, but in this dataset, it appears not to be a major factor.
2. **Age**: The Hazard Ratio (HR) for Age is higher than 1 (1.103996) and highly significant (p-value = 3.82e-07), indicating that for each additional year of age, the risk of death increases by about 10.39%. Age is a known risk factor in survival studies, particularly in cardiac procedures, where older patients typically have a higher risk due to greater frailty and the likelihood of comorbid conditions.
3. **Coronary Artery Bypass Grafting (CABG) (con.cabg)**: The presence of CABG (indicated by con.cabgYes) is associated with a significantly higher risk of death (p-value = 0.027932). The hazard ratio of 1.916 suggests that patients who have undergone CABG have about 91.6% higher risk of death compared to those who have not. This might reflect the severity of coronary artery disease in patients requiring CABG, which can complicate heart valve surgery outcomes.
4. Creatinine Levels (creat): The creatinine level shows a minimal impact on the hazard (coef = 0.006730) with a p-value of 0.131411, suggesting that within the range of creatinine levels in this study population, there isn't a strong or significant association with survival outcomes. Creatinine levels are a marker of kidney function, which can be crucial in overall health status but may not directly influence survival outcomes post-surgery unless reaching more extreme values.
5. **Left Ventricular Function (lv)**: Given a level of significance equal to 0.05, a moderate ventricular function (lvModerate) seems to not be a significant predictor for mortality following heart valve transplant. On the contrary, lvLow is highly significant (p-value = 0.000232) with a Hazard Ratio of 5.182, indicating a substantial increase (more than 5 times more) in risk for patients with low ventricular function. This is consistent with clinical expectations, as poor ventricular function can severely impact the heart's ability to pump blood effectively after surgery.
6. Stenosis Severity (sten.reg.mix): The severity of valve stenosis or mixed conditions shows no significant effect on survival at the 0.05 level of significance. However, the negative coefficients for regurgitation (-0.619062) and mixed (-0.812285) suggest a protective effect compared to stenosis (though, again, not statistically significant). This might indicate different pathophysiological impacts of these conditions on patient outcomes, potentially moderated by other factors not included in the model.

Overall, the base model suggests that age, coronary artery bypass grafting, left ventricular function and stenosis severity are significant predictors of survival following heart valve transplant. Sex and creatinine levels (creat) do not appear to significantly affect survival. 

The model has moderate predictive accuracy (Concordance = 0.783) and overall good fit according to model fit tests.


To evaluate the functional form of continuous variables and ensure that the proportional hazards assumption holds for each covariate, we will visually inspect the Schoenfeld Residuals. Moreover, we will check the linearity of each continuous variable against the null model (without covariates).

```{r Schoenfeld and Martingale Residuals}
# Evaluate the proportional hazards assumption using cox.zph function
# This function performs a test of proportional hazards based on the scaled Schoenfeld residuals
ph_test <- cox.zph(base_model)

# Plotting the residuals to visually inspect any trend indicating non-proportionality
plot(ph_test)  # Multiple plots if more than one covariate; assess each plot for trends over time

# Output the global test of proportional hazards from cox.zph
print(ph_test) # This will display the test statistics and p-values for each covariate

# Check the linearity of continuous variables using Martingale residuals from a null model
null_model <- coxph(Surv(fuyrs, status) ~ 1, data = data) # Null model with no covariates

# Selecting continuous variables to check for non-linearity
continuous_vars <- c("age", "creat")

# Plot Martingale residuals against each continuous variable
for(var in continuous_vars) {
  # Plot residuals against continuous variables
  plot(data[[var]], residuals(null_model, type = "martingale"), xlab = var, ylab = "Martingale residuals", main = paste("Linearity Check for", var))
  abline(h = 0, col = "red")  # Horizontal line at zero to aid in visual assessment
}

```


## Augmented Model

```{r Augmented Model}
augmented_model <- coxph(Surv(fuyrs, status) ~ sex + age + con.cabg + creat + lv + sten.reg.mix + log.lvmi, data = data, x=T)

summary(augmented_model)
```


The results from the new augmented model, which now includes *log.lvmi* alongside other variables, provide significant insights into the factors influencing survival after heart valve implantation. In particular:
1. Sex (Female): While the coefficient for females has increased slightly compared to the base model, it remains non-significant (p = 0.225493). The hazard ratio is 1.486, indicating a 48.6% higher risk of death for females compared to males, though this difference is not statistically robust. This might suggest some inherent but not definitive differences in outcomes by gender that are not captured fully by the model or sample size.
2. **Age**: Consistently significant across both models, each additional year of age increases the risk of death by 10.3% (p = 6.15e-07). This confirms the substantial impact of age on survival post-surgery, reflecting the increased vulnerability and potential comorbidities associated with older age.
3. **Coronary Artery Bypass Grafting (CABG) (Yes)**: The hazard ratio for CABG has slightly increased to 2.090, and it remains significant (p = 0.014313). This underlines the higher risk associated with patients who have had CABG, likely reflecting more severe baseline coronary artery disease.
4. Creatinine (Creat): The effect of creatinine levels remains non-significant (p = 0.459702) with a very minimal hazard ratio change (1.004). This suggests that within this study population, creatinine levels are not a strong predictor of survival, possibly due to not reaching clinically critical levels.
5. **Left Ventricular Function (LV)**: The effect of moderate ventricular function, in this augmented model, seems to become statistically significant at a level of 0.05 (p-value=0.047024), while the low ventricular function still remains highly significant (p-value=0.000189) with a very high hazard ratio of 5.272097. Again, low ventricular function is a strong predictor for both models, representing a significant risk factor for mortality post-surgery.
6. **Stenosis Severity (Sten.reg.mix)**: The effect of stenosis severity remains non-significant for regurgitation (p-value=0.150625), while becoming statistically significant for mixed conditions (p-value=0.037195). Compared to stenosis, mixed conditions seem to show a protective effect (HR=0.392434, or 39.2% lower risk).
7. **Logarithm of Left Ventricular Mass Index (log.lvmi)**: This newly added variable is significant (p = 0.001712) with a coefficient of 1.219381, translating into a hazard ratio of 3.385. This indicates that higher values of log-transformed left ventricular mass index substantially increase the risk of death.

The inclusion of this last variable not only highlights its importance in clincical effects but also enhances the model's explanatory power, as seen in the improved concordance and significant test statistics (Likelihood ratio, Wald test, and logrank test).

Again, to evaluate the functional form of continuous variables and ensure that the proportional hazards assumption holds for each covariate, we will visually inspect the Schoenfeld Residuals. Moreover, we will check the linearity of each continuous variable against the null model (without covariates).

```{r Schoenfeld and Martingale Residuals}
# Evaluate the proportional hazards assumption using cox.zph function
# This function performs a test of proportional hazards based on the scaled Schoenfeld residuals
ph_test_augmented <- cox.zph(augmented_model)

# Plotting the residuals to visually inspect any trend indicating non-proportionality
plot(ph_test_augmented)  # Multiple plots if more than one covariate; assess each plot for trends over time
abline(h = 0, col = "red") 
# Output the global test of proportional hazards from cox.zph
print(ph_test_augmented) # This will display the test statistics and p-values for each covariate

# Selecting continuous variables to check for non-linearity
continuous_vars_augmented <- c("age", "creat", "log.lvmi")  # Now including log.lvmi

# Plot Martingale residuals against each continuous variable
for(var in continuous_vars_augmented) {
  # Plot residuals against continuous variables
  plot(data[[var]], residuals(null_model, type = "martingale"), xlab = var, ylab = "Martingale residuals", main = paste("Linearity Check for", var))
  abline(h = 0, col = "red")  # Horizontal line at zero to aid in visual assessment
}
```



## Augmented Model - Polynomial Transformation for PH

```{r Augmented Polynomial}
augmented_model_poly <- coxph(Surv(fuyrs, status) ~ sex + age + con.cabg + creat + lv + sten.reg.mix + poly(log.lvmi, 3), data = data, x=T)

summary(augmented_model_poly)
```

```{r Schonefeld and Martingale Residuals}
# Evaluate the proportional hazards assumption using the cox.zph function
# This function tests the proportional hazards assumption based on the scaled Schoenfeld residuals
ph_test_2 <- cox.zph(augmented_model_poly)

# Plotting the residuals to visually inspect any trend indicating non-proportionality
plot(ph_test_2)  # This generates multiple plots if there are multiple covariates

# Print the global test of proportional hazards
print(ph_test_2) # Displays test statistics and p-values for each covariate
```
## Augmented Model - Stratified

```{r}
# Categorize log.lvmi into tertiles or quartiles
data$log.lvmi_cat <- cut(data$log.lvmi, breaks = quantile(data$log.lvmi, probs = seq(0, 1, by = 0.25)), include.lowest = TRUE)

# Fit the Cox model with stratification by log.lvmi categories
stratified_model <- coxph(Surv(fuyrs, status) ~ sex + age + con.cabg + creat + lv + sten.reg.mix + strata(log.lvmi_cat), data = data)
summary(stratified_model)

# Check proportional hazards assumption
cox.zph_stratified <- cox.zph(stratified_model)
print(cox.zph_stratified)


```
The strata function in the coxph model does not include the stratified variable (log.lvmi_cat) in the coefficients table because stratification means the model estimates separate baseline hazard functions for each stratum without directly estimating coefficients for the stratified variable. This allows the hazard functions to differ across the strata, addressing the proportional hazards assumption for the stratified variable without explicitly modeling its effect as a coefficient.


The results indicate that after stratifying by log.lvmi_cat, the proportional hazards assumption is satisfied for all covariates (as seen from the cox.zph test results).

```{r}
augmented_model_str <- coxph(Surv(fuyrs, status) ~ sex + age + con.cabg + creat + lv + sten.reg.mix + strata(log.lvmi), data = data, x=T)

summary(augmented_model_str)
```


```{r}
ph_test_strata <- cox.zph(augmented_model_strata)

plot(ph_test_strata)  # This generates multiple plots if there are multiple covariates

# Print the global test of proportional hazards
print(ph_test_strata)
```




## Augmented Model - Time Varying

```{r}
augmented_model <- coxph(Surv(fuyrs, status) ~ sex + age + con.cabg + creat + lv + sten.reg.mix + log.lvmi, data = data, x=T)

summary(augmented_model)

cox.zph_res <- cox.zph(augmented_model)
plot(cox.zph_res, var = "log.lvmi")
```

```{r}

data$log.lvmi_time <- data$log.lvmi * data$fuyrs
time_dep_model <- coxph(Surv(fuyrs, status) ~ sex + age + con.cabg + creat + lv + sten.reg.mix + log.lvmi_time, data = data, x=T)
summary(time_dep_model)
```

```{r}
cox.zph_time_dep <- cox.zph(time_dep_model)
print(cox.zph_time_dep)

```


```{r Schonefeld and Martingale Residuals}

martingale_residuals <- residuals(time_dep_model, type = "martingale")
plot(martingale_residuals, ylab = "Martingale Residuals", xlab = "Observation Index", main = "Martingale Residuals Plot")
abline(h = 0, col = "red")

# Check Schoenfeld residuals
schoenfeld_test <- cox.zph(time_dep_model)
print(schoenfeld_test)
plot(schoenfeld_test)
```
It seems that including the time-dependent term for log.lvmi (log.lvmi_time) introduced new violations of the proportional hazards assumption for other covariates. This indicates that adding the interaction term may not be the appropriate way to handle the time-varying effect of log.lvmi


# Performance Evaluation

## Calibration

```{r Base Model predictions}
fit1<-survfit(base_model,newdata=data)
data$risk_base<-1-as.numeric(summary(fit1,times=5)$surv)
```

```{r Augmented Model predictions}
fit2<-survfit(augmented_model,newdata=data)
data$risk_augmented<-1-as.numeric(summary(fit2,times=5)$surv)
```

```{r Augmented Model Poly predictions}
fit3<-survfit(augmented_model_poly,newdata=data)
data$risk_augmented_poly<-1-as.numeric(summary(fit3,times=5)$surv)
```


```{r Augmented Model Strata predictions}
fit4<-survfit(stratified_model,newdata=data)
data$risk_augmented_strata<-1-as.numeric(summary(fit4,times=5)$surv)
```



```{r Scores for ROC and Calibration Plot}
score <- Score(list("Base"=base_model,"Augmented"=augmented_model, "Poly"=augmented_model_poly),
              formula=Surv(fuyrs, status==1)~1,
              data=data,conf.int=T,
              times=seq(1,5,1),
              plots=c("calibration","ROC"))

plotCalibration(score,times=5,cens.method="local",method="quantile",q=10)
title(main="Calibration Plot at 5 years")

```
```{r Plotting the ROC Curves}
plotROC(score,times=5,cens.method="local")
title(main="time-dependent ROC at 5 years")
```

```{r Bootstrapping for internal validation}
CVscore <- Score(list("Base"=base_model,"Augmented"=augmented_model,  "Poly"=augmented_model_poly),
              formula=Surv(fuyrs, status==1)~1,
              data=data,conf.int=T,
              times=seq(1,5,1),
              split.method="loob",B=100,seed=1000)

summary(CVscore)
```


```{r}
library(survival)
library(survminer)

# Fit the models
fit1 <- survfit(base_model, newdata=data)
fit2 <- survfit(augmented_model, newdata=data)
fit3 <- survfit(augmented_model_poly, newdata=data)
fit4 <- survfit(stratified_model, newdata=data)

# Calculate risk for each model
data$risk_base <- 1 - as.numeric(summary(fit1, times=5)$surv)
data$risk_augmented <- 1 - as.numeric(summary(fit2, times=5)$surv)
data$risk_augmented_poly <- 1 - as.numeric(summary(fit3, times=5)$surv)
data$risk_augmented_strata <- 1 - as.numeric(summary(fit4, times=5)$surv)

# Calculate Concordance Index (C-index)
cindex_base <- summary(base_model)$concordance
cindex_augmented <- summary(augmented_model)$concordance
cindex_augmented_poly <- summary(augmented_model_poly)$concordance
cindex_stratified <- summary(stratified_model)$concordance

cindex_base
cindex_augmented
cindex_augmented_poly
cindex_stratified

# Plot Kaplan-Meier Curves
#ggsurvplot(fit1, data = data, risk.table = TRUE, title = "Base Model")
#ggsurvplot(fit2, data = data, risk.table = TRUE, title = "Augmented Model")
#ggsurvplot(fit3, data = data, risk.table = TRUE, title = "Augmented Model with Polynomial Terms")
#ggsurvplot(fit4, data = data, risk.table = TRUE, title = "Stratified Model")

# Perform Likelihood Ratio Test
anova(base_model, augmented_model, augmented_model_poly, stratified_model, test="LRT")

# Calculate AIC
aic_base <- AIC(base_model)
aic_augmented <- AIC(augmented_model)
aic_augmented_poly <- AIC(augmented_model_poly)
aic_stratified <- AIC(stratified_model)

aic_base
aic_augmented
aic_augmented_poly
aic_stratified

```
The C-index values are relatively close to each other, indicating that all models have a similar discriminatory power. The augmented models (with and without polynomial terms) have slightly higher C-index values compared to the base model, suggesting a modest improvement in predictive accuracy. The stratified model also shows a high C-index but is slightly lower than the augmented models.

The likelihood ratio test results show that the augmented model (Model 2) significantly improves the fit over the base model (Model 1) with a p-value of 0.001966. However, adding polynomial terms (Model 3) does not significantly improve the fit compared to the augmented model (Model 2), as indicated by the non-significant p-value of 0.375014.

The stratified model (Model 4) shows a substantial improvement over the other models, with a highly significant p-value (p < 2.2e-16). This suggests that incorporating stratification based on log.lvmi_cat significantly improves the model fit.

The AIC values indicate that the stratified model (Model 4) has the lowest AIC, suggesting it provides the best fit among the models. The augmented model (Model 2) has a lower AIC than the base model (Model 1), indicating a better fit. The augmented model with polynomial terms (Model 3) has a slightly higher AIC than Model 2, supporting the likelihood ratio test result that adding polynomial terms does not significantly improve the model.

In conclusion, the stratified model (Model 4) is the best-performing model, providing significant improvements in fit and predictive accuracy. The simple augmented model (Model 2) also performs well, offering a good balance of complexity and performance.

### Brier Score
```{r}
# indicator of death within 5 year
data$status.5y<-ifelse(data$fuyrs<=5,1,0)
# Brier Score
(BS_base <- mean((data$status.5y - data$risk_base) ^ 2)) 
(BS_augmented <- mean((data$status.5y - data$risk_augmented) ^ 2))
(BS_poly <- mean((data$status.5y - data$risk_augmented_poly) ^ 2))
(BS_strata <- mean((data$status.5y - data$risk_augmented_strata) ^ 2)) 

```
The Brier scores for your models provide insight into their predictive accuracy at the 5-year mark. The base model has the lowest Brier score (0.3476), indicating the best predictive performance. The augmented model (0.3550) and the stratified model (0.3547) both have slightly higher Brier scores, suggesting a marginal decrease in predictive accuracy compared to the base model. The augmented model with polynomial terms has the highest Brier score (0.3593), indicating that the added complexity did not improve and actually slightly worsened predictive accuracy. Overall, the base model performs the best in terms of predictive accuracy, while the augmented and stratified models do not offer substantial improvements and the polynomial terms appear to detract from the model's performance.

The discrepancy between the Brier scores and metrics like AIC and the likelihood ratio test arises because they measure different aspects of model performance. The Brier score directly assesses prediction accuracy at a specific time point (5 years), while AIC evaluates overall model fit with a penalty for complexity, and the likelihood ratio test checks for significant improvements with additional predictors.

In your case, the base model has the best Brier score, indicating it predicts 5-year survival most accurately. However, the augmented and stratified models have lower AIC values and significant improvements in the likelihood ratio test, suggesting better overall fit to the training data but possibly overfitting, leading to less accurate predictions on new data. The augmented model with polynomial terms, while fitting the training data well, introduces complexity that worsens its predictive accuracy.


```{r}
#????? Calculation of Model Improvement over Random Guessing Using Brier Scores
BS_basesc  <- mean(data$risk_base*(1-data$risk_base))
BS_augmentedsc <- mean(data$risk_augmented*(1-data$risk_augmented))

BS_base-BS_basesc
BS_augmented-BS_augmentedsc
```

### ROC AUC
```{r}

###TO DELETE
roc1<-roc(data$status.5y, data$risk_base)
plot(1 - roc1$specificities, roc1$sensitivities, 
     type = 'l', ylab = 'TPF', xlab = 'FPF', lwd = 3, xaxt = "n", yaxt = "n", 
     xlim = c(0,1), cex.lab = 1.7, frame = F)
axis(1, at = c(0,0.25,0.5,0.75,1), labels = NA, pos = 0)
axis(1, at = c(0,0.25,0.5,0.75,1), labels = c(0,0.25,0.5,0.75,1), cex.axis = 1.7, pos = 0)
axis(2, at = c(0,0.25,0.5,0.75,1), labels = c(0,0.25,0.5,0.75,1), cex.axis = 1.7, pos = 0)
Youden1<-roc1$sensitivities+roc1$specificities-1
optimal.cut.off1<-roc1$thresholds[Youden1==max(Youden1)]
cbind(optimal.cut.off1,Youden=max(Youden1))
points(1-roc1$specificities[roc1$thresholds==optimal.cut.off1],roc1$sensitivities[roc1$thresholds==optimal.cut.off1],pch=0,cex=1.7)

roc2<-roc(data$status.5y, data$risk_augmented)

lines(1 - roc2$specificities, roc2$sensitivities, 
      lwd = 3, lty = 3)
Youden2<-roc2$sensitivities+roc2$specificities-1
optimal.cut.off2<-roc2$thresholds[Youden2==max(Youden2)]
cbind(optimal.cut.off2,Youden=max(Youden2))


points(1-roc2$specificities[roc2$thresholds==optimal.cut.off2],roc2$sensitivities[roc2$thresholds==optimal.cut.off2],pch=0,cex=1.7)

# Chance line:
abline(a=0, b=1, lty = 2, lwd = 2)
lines(c(0,1), c(1,1), lty = 1)
lines(c(1,1), c(0,1), lty = 1)

legend(x = 0, y = 1, c("model1","model2"), lwd = c(3,3), lty = c(1,3), bty = 'n', cex = 1.7)


```




```{r}
#??????
(AUC1 <- roc1$auc)
(AUC2 <- roc2$auc)   
AUC2 - AUC1
```
###NET BENEFIT

```{r}
library(dcurves)
dca(Surv(fuyrs, status==1) ~ risk_base + risk_augmented + risk_augmented_poly + risk_augmented_strata, data, time = 5)
```
# Model Predictions



##TEST



```{r}

# Prediction of event risk at a fixed time-point (e.g., 5 years) for 3 "type" subjects
# Create example subjects
type_subjects <- data.frame(
  sex = c("Male", "Female", "Male"),
  age = c(100, 70, 65),
  con.cabg = c("Yes", "No", "Yes"),
  creat = c(100, 150, 120),
  lv = c("High", "Moderate", "Low"),
  sten.reg.mix = c("Stenosis", "Regurgitation", "Mixed"),
  log.lvmi = c(log(150), log(170), log(160))
)

# Predict event risk at 5 years using both models
risk_basic <- predict(base_model, newdata = type_subjects, type = "risk", se.fit = TRUE)
risk_augmented_2 <- predict(augmented_model_poly, newdata = type_subjects, type = "risk", se.fit = TRUE)

risk_basic
risk_augmented_2

```
```{r}


# Create example subjects
type_subjects <- data.frame(
  sex = c("Male", "Male", "Male"),
  age = c(70, 80, 70),
  con.cabg = c("No", "Yes", "Yes"),
  creat = c(130, 150, 130),
  lv = c("Moderate", "High", "Moderate"),
  sten.reg.mix = c("Mixed", "Mixed", "Mixed"),
  log.lvmi = c(log(160), log(200), log(160))
)

# Predict survival probabilities at 5 years for base model
surv_base <- survfit(base_model, newdata = type_subjects)
risk_base_5yr <- 1-summary(surv_base, times = 5)$surv
risk_base_5yr


# Predict survival probabilities at 5 years for augmented model
surv_augmented <- survfit(augmented_model_poly, newdata = type_subjects)
risk_augmented_5yr <- 1-summary(surv_augmented, times = 5)$surv
risk_augmented_5yr
```

