---
title: "data.data"
output:
  html_document:
    df_print: paged
editor_options: 
  markdown: 
    wrap: sentence
---

**Project Big Data in Healthcare**

*Alessandro Rota(798050), Matteo Mondini(902873), Alexandre Crivellari(902064), Yasmin Bouhdada(837389).*

# Index

1.  [Library]

2.  [Importing Data]

3.  [Dataset Variables]

4.  [Descriptive analysis]

    1.  [Dataset info]

    2.  [Age analysis]

    3.  [Sex analysis]

    4.  [Correlation between variables]

    5.  [Detailed Variable Description]

5.  [Models]

    1.  [Univariate Models]

    2.  [Base model]

    3.  [Augmented Model]

    4.  [Augmented Model - Polynomial Transformation for PH]

6.  [Performance Evaluation]

    1.  [Calibration]
        -   [Brier Score]
        -   [ROC AUC]
        -   [NET BENEFIT]

7.  [Model Predictions]


# Libraries

```{r Libraries}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2) # For plotting
library(dplyr) # For data handling
library(readr)
library(survival) # Main library, used for implementing Cox Modeling
library(pROC) # For plotting ROC curves
library(survminer)
library(utils) #For not having to upload the txt each time
library(riskRegression) # For performance indicators
```

# Importing Data

```{r Importing data from shared GitHub}

#setwd('~/Desktop/Data Science/Data Science Lab in Medicine/Big Data in Health Care/')
url <- "https://raw.githubusercontent.com/darthgween/BigDataInHealthCareProject/main/heart.valve.txt?token=GHSAT0AAAAAACR3GYSXO5Z3ARKVQY2S6SYSZRYZQWA"
data <- read.table(url, header = TRUE)
head(data)
```

```{r Transforming variables into Categoric}

data$sex <- factor(data$sex, levels = c(0, 1), labels = c("Male", "Female"))
data$con.cabg <- factor(data$con.cabg, levels = c(0, 1), labels = c("No", "Yes"))
data$lv <- factor(data$lv, levels = c(1, 2, 3), labels = c("High", "Moderate", "Low"))
data$sten.reg.mix <- factor(data$sten.reg.mix, levels = c(1, 2, 3), labels = c("Stenosis", "Regurgitation", "Mixed"))

```

# Descriptive analysis

## Description of dataset variables:

-   ***Paz.id***: patient identification number.

-   ***log.lvmi***: natural logarithm of the “Left Ventricular Mass Index” (standardized) measured at baseline.

-   ***fuyrs***: follow-up time from surgery, in years.

-   ***status***: event indicator (1 = dead; 0 = censored).

-   ***sex***: sex of the patient (0 = M; 1 = F).

-   ***age***: age of the patient (years) at surgery.

-   ***con.cabg***: presence of coronary bypass (1 = yes; 0 = no).

-   ***creat***: pre-operative serum creatinine (µmol/mL).

-   ***lv***: pre-operative left ventricular ejection fraction (1 = high, 2 = moderate, 3 = low).

-   ***sten.reg.mix***: hemodynamics of the heart valve (1 = stenosis, 2 = regurgitation, 3 = mixed).

### log.lvmi: Natural logarithm of the “Left Ventricular Mass Index” (standardized) measured at baseline

The Left Ventricular Mass Index (LVMI) is a measurement that reflects the mass of the left ventricle of the heart relative to the individual’s body surface area (BSA).
It is an important indicator of cardiac health, often used to diagnose and monitor conditions like left ventricular hypertrophy (LVH), which is the thickening of the heart’s left ventricular walls.

**Normal and Balanced Values:**

Men:

• Normal LVMI: less than 115 g/m²

• Logarithmic (ln) Normal LVMI: ln(115) ≈ 4.74

Women:

• Normal LVMI: less than 95 g/m²

• Logarithmic (ln) Normal LVMI: ln(95) ≈ 4.55

```{r Boxplot log.lvmi - Status}
ggplot(data, aes(x = factor(status), y = log.lvmi, fill = factor(status))) +
  geom_boxplot() +
  ggtitle("Boxplot of log.lvmi for Status") +
  xlab("Status") +
  ylab("log.lvmi") +
  scale_fill_discrete(name = "Status")

```

### creat: Pre-operative serum creatinine levels (µmol/L)

A measure of kidney function.
Normal levels are 44-80 µmol/L for women and 62-106 µmol/L for men.
Elevated creatinine levels suggest renal impairment, a risk factor for surgical complications.

### lv: Pre-operative left ventricular ejection fraction

A measure of how well the left ventricle pumps blood with each beat.Low ejection fraction indicates heart failure and a higher surgical risk

### sten.reg.mix: hemodynamics of the heart valve (1 = stenosis, 2 = regurgitation, 3 = mixed)

Describes the hemodynamic characteristics of the heart valve.
Different types of valve disease require specific surgical and medical management.

## Dataset Information

- Dimensions

```{r Dimensions}
dim(data)
```

- Object structure

```{r Check for data structure and types}
str(data)
```
- Data Summary

```{r Dataset Summary for all variables}
summary(data)
```

## Age distribution

This information is crucial for understanding the demographics of the patient population and tailoring perioperative care.
Typically, heart valve disease prevalence increases with age, with most patients being in the older age brackets.
For instance, older patients (above 65 years) are at higher risk for complications and may require more intensive postoperative care.

```{r Age Histogram}
ggplot(data, aes(x = age)) +
  geom_histogram(binwidth = 5, fill = "lightblue", color = "black") +
  ggtitle("Age Distribution") +
  xlab("Age") +
  ylab("Frequency")
```

## Gender distribution

Out of 255 Observations 73 of them are Female and 182 are male.

```{r}
sum_males <- sum(data$sex == "Male")
sum_females <- sum(data$sex == "Female")

cat("Number of males:", sum_males, "\n")
cat("Number of females:", sum_females, "\n")
```

- Bar plot

```{r}
ggplot(data, aes(x = sex, fill = sex)) + 
  geom_bar() + 
  scale_fill_manual(values = c("Male" = "blue", "Female" = "pink")) +
  xlab("Sex") + 
  ylab("Count") +
  ggtitle("Distribution by Sex") +
  theme_minimal()

```

## Correlation between variables

```{r Correlation Matrix}
#cor_matrix <- cor(data)
#print(cor_matrix)

#if (!require(corrplot)) {
    install.packages("corrplot")
    library(corrplot)
#}
#corrplot(cor_matrix, method = "circle")
```


```{r Scatterpot log.lvmi - Age}
ggplot(data, aes(x = age, y = log.lvmi)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  ggtitle("Scatter Plot: età vs log.lvmi") +
  xlab("Età") +
  ylab("log.lvmi")

```

# Models

## Univariate Models
Here we cycle through all variables of the dataset, excluding: the *patient ID* variable (which cannot possibly explain anything about the survival probability of patients), the *status* variable (which is the outcome) and the *fuyrs* variable (as it describes the follow-up time - hence, the censoring time).


```{r Univariate Function loop}
# The outcome variable is a survival object created with the time to event (fuyrs) and status indicator (status)
surv_obj <- Surv(data$fuyrs, data$status)

# Perform Cox regression for each predictor
# Excluding paz.id from the analysis as it's an identifier, not a predictor
variables <- c("log.lvmi", "sex", "age", "con.cabg", "creat", "lv", "sten.reg.mix")

# Store results in a list for easy access
univariate_models <- list()

for (var in variables) {
  # Formula to dynamically create the regression model
  formula <- as.formula(paste("surv_obj ~", var))
  
  # Fit the Cox model
  model <- coxph(formula, data = data)
  
  # Print the summary of the model
  print(summary(model))
  
  # Store the model in the list
  univariate_models[[var]] <- model
}
```
Below list of R code snippets is provided, in order to easily access each Univariate model, independently.

```{r Univariate Left Ventricular Mass Index}
univariate_models[["log.lvmi"]]
```

```{r Univariate Sex}
univariate_models[["sex"]]
```

```{r Univariate Age}
univariate_models[["age"]]
```

```{r Univariate Coronary Bypass}
univariate_models[["con.cabg"]]
```


```{r Univariate Creatinine}
univariate_models[["creat"]]
```


```{r Univariate Left Ventricular Ejection Fraction}
univariate_models[["lv"]]
```


```{r Univariate Hemodynamics of Heart Valve}
univariate_models[["sten.reg.mix"]]
```
## Base model

```{r Base Model}
base_model <- coxph(Surv(fuyrs, status) ~ sex + age + con.cabg + creat + lv + sten.reg.mix, data = data, x=T)

summary(base_model)
```

The Cox proportional hazards model shows significant effects of age, coronary artery bypass grafting (con.cabg), left ventricular function (lv), and stenosis severity (sten.reg.mix) on survival time. In particular:
1. Sex: the variable sex does not show a significant effect (p-value = 0.392228) on survival. The hazard ratio for females (1.317) would suggest a higher risk compared to males, but the wide confidence interval (0.7009 to 2.475) and high p-value indicate that this effect is not statistically robust. In many clinical studies, sex may influence outcomes due to differences in biological responses or treatment approaches, but in this dataset, it appears not to be a major factor.
2. **Age**: The Hazard Ratio (HR) for Age is higher than 1 (1.103996) and highly significant (p-value = 3.82e-07), indicating that for each additional year of age, the risk of death increases by about 10.39%. Age is a known risk factor in survival studies, particularly in cardiac procedures, where older patients typically have a higher risk due to greater frailty and the likelihood of comorbid conditions.
3. **Coronary Artery Bypass Grafting (CABG) (con.cabg)**: The presence of CABG (indicated by con.cabgYes) is associated with a significantly higher risk of death (p-value = 0.027932). The hazard ratio of 1.916 suggests that patients who have undergone CABG have about 91.6% higher risk of death compared to those who have not. This might reflect the severity of coronary artery disease in patients requiring CABG, which can complicate heart valve surgery outcomes.
4. Creatinine Levels (creat): The creatinine level shows a minimal impact on the hazard (coef = 0.006730) with a p-value of 0.131411, suggesting that within the range of creatinine levels in this study population, there isn't a strong or significant association with survival outcomes. Creatinine levels are a marker of kidney function, which can be crucial in overall health status but may not directly influence survival outcomes post-surgery unless reaching more extreme values.
5. **Left Ventricular Function (lv)**: Given a level of significance equal to 0.05, a moderate ventricular function (lvModerate) seems to not be a significant predictor for mortality following heart valve transplant. On the contrary, lvLow is highly significant (p-value = 0.000232) with a Hazard Ratio of 5.182, indicating a substantial increase (more than 5 times more) in risk for patients with low ventricular function. This is consistent with clinical expectations, as poor ventricular function can severely impact the heart's ability to pump blood effectively after surgery.
6. Stenosis Severity (sten.reg.mix): The severity of valve stenosis or mixed conditions shows no significant effect on survival at the 0.05 level of significance. However, the negative coefficients for regurgitation (-0.619062) and mixed (-0.812285) suggest a protective effect compared to stenosis (though, again, not statistically significant). This might indicate different pathophysiological impacts of these conditions on patient outcomes, potentially moderated by other factors not included in the model.

Overall, the base model suggests that age, coronary artery bypass grafting, left ventricular function and stenosis severity are significant predictors of survival following heart valve transplant. Sex and creatinine levels (creat) do not appear to significantly affect survival. 

The model has moderate predictive accuracy (Concordance = 0.783) and overall good fit according to model fit tests.

To evaluate the functional form of continuous variables and ensure that the proportional hazards assumption holds for each covariate, we will visually inspect the Schoenfeld Residuals. Moreover, we will check the linearity of each continuous variable against the null model (without covariates).

```{r Schoenfeld and Martingale Residuals}
# Evaluate the proportional hazards assumption using cox.zph function
# This function performs a test of proportional hazards based on the scaled Schoenfeld residuals
ph_test <- cox.zph(base_model)

# Plotting the residuals to visually inspect any trend indicating non-proportionality
plot(ph_test)  # Multiple plots if more than one covariate; assess each plot for trends over time

# Output the global test of proportional hazards from cox.zph
print(ph_test) # This will display the test statistics and p-values for each covariate

# Check the linearity of continuous variables using Martingale residuals from a null model
null_model <- coxph(Surv(fuyrs, status) ~ 1, data = data) # Null model with no covariates

# Selecting continuous variables to check for non-linearity
continuous_vars <- c("age", "creat")

# Plot Martingale residuals against each continuous variable
for(var in continuous_vars) {
  # Plot residuals against continuous variables
  plot(data[[var]], residuals(null_model, type = "martingale"), xlab = var, ylab = "Martingale residuals", main = paste("Linearity Check for", var))
  abline(h = 0, col = "red")  # Horizontal line at zero to aid in visual assessment
}

```
The cox.zph function was used to compute the Schoenfeld and Martingale residuals, which are needed for testing the proportional hazards and linearity of the continuous variable assumptions. Each covariate's residuals are tested to see if they are  independent against time (which they should be if the proportional hazards assumption holds true).
The output provided indicates, both graphically and from the p-value table, that no covariate show evidence against the proportional hazards assumptions. The GLOBAL result, with a p-value of 0.84, suggests that the covariates satisfy (overall) the proportional hazards assumption of the model.

## Augmented Model

```{r Augmented Model}
augmented_model <- coxph(Surv(fuyrs, status) ~ sex + age + con.cabg + creat + lv + sten.reg.mix + log.lvmi, data = data, x=T)

summary(augmented_model)
```


The results from the new augmented model, which now includes *log.lvmi* alongside other variables, provide significant insights into the factors influencing survival after heart valve implantation. In particular:
1. Sex (Female): While the coefficient for females has increased slightly compared to the base model, it remains non-significant (p = 0.225493). The hazard ratio is 1.486, indicating a 48.6% higher risk of death for females compared to males, though this difference is not statistically robust. This might suggest some inherent but not definitive differences in outcomes by gender that are not captured fully by the model or sample size.
2. **Age**: Consistently significant across both models, each additional year of age increases the risk of death by 10.3% (p = 6.15e-07). This confirms the substantial impact of age on survival post-surgery, reflecting the increased vulnerability and potential comorbidities associated with older age.
3. **Coronary Artery Bypass Grafting (CABG) (Yes)**: The hazard ratio for CABG has slightly increased to 2.090, and it remains significant (p = 0.014313). This underlines the higher risk associated with patients who have had CABG, likely reflecting more severe baseline coronary artery disease.
4. Creatinine (Creat): The effect of creatinine levels remains non-significant (p = 0.459702) with a very minimal hazard ratio change (1.004). This suggests that within this study population, creatinine levels are not a strong predictor of survival, possibly due to not reaching clinically critical levels.
5. **Left Ventricular Function (LV)**: The effect of moderate ventricular function, in this augmented model, seems to become statistically significant at a level of 0.05 (p-value=0.047024), while the low ventricular function still remains highly significant (p-value=0.000189) with a very high hazard ratio of 5.272097. Again, low ventricular function is a strong predictor for both models, representing a significant risk factor for mortality post-surgery.
6. **Stenosis Severity (Sten.reg.mix)**: The effect of stenosis severity remains non-significant for regurgitation (p-value=0.150625), while becoming statistically significant for mixed conditions (p-value=0.037195). Compared to stenosis, mixed conditions seem to show a protective effect (HR=0.392434, or 39.2% lower risk).
7. **Logarithm of Left Ventricular Mass Index (log.lvmi)**: This newly added variable is significant (p = 0.001712) with a coefficient of 1.219381, translating into a hazard ratio of 3.385. This indicates that higher values of log-transformed left ventricular mass index substantially increase the risk of death.

The inclusion of this last variable not only highlights its importance in clincical effects but also enhances the model's explanatory power, as seen in the improved concordance and significant test statistics (Likelihood ratio, Wald test, and logrank test).

Again, to evaluate the functional form of continuous variables and ensure that the proportional hazards assumption holds for each covariate, we will visually inspect the Schoenfeld Residuals. Moreover, we will check the linearity of each continuous variable against the null model (without covariates).

```{r Schoenfeld and Martingale Residuals}
# Evaluate the proportional hazards assumption using cox.zph function
# This function performs a test of proportional hazards based on the scaled Schoenfeld residuals
ph_test_augmented <- cox.zph(augmented_model)

# Plotting the residuals to visually inspect any trend indicating non-proportionality
plot(ph_test_augmented)  # Multiple plots if more than one covariate; assess each plot for trends over time

# Output the global test of proportional hazards from cox.zph
print(ph_test_augmented) # This will display the test statistics and p-values for each covariate

# Selecting continuous variables to check for non-linearity
continuous_vars_augmented <- c("age", "creat", "log.lvmi")  # Now including log.lvmi

# Plot Martingale residuals against each continuous variable
for(var in continuous_vars_augmented) {
  # Plot residuals against continuous variables
  plot(data[[var]], residuals(null_model, type = "martingale"), xlab = var, ylab = "Martingale residuals", main = paste("Linearity Check for", var))
  abline(h = 0, col = "red")  # Horizontal line at zero to aid in visual assessment
}
```
In this case, the graphical result for the newly introduced log.lvmi variable show a *plausible* violation of the Proportional Hazards assumption, with a slight increasing trend of the coefficient as a dependance of time and another slight visual concavity. The p-value of the test (equal to 0.012) would suggest to interpret the Proportional Hazards assumption as invalid, although the graphical trend would not seem too definitive to rule it out too strongly.

In this situation of uncertainty, we have decided to implement a *stratification* of the model using the Strata function for log.lvmi, in order to obtain an additional model that would more strongly maintain the Proportional Hazards assumption for the interpretability of the results.

## Augmented Model - Stratified

```{r Augmented - Stratified Model}
# Categorize log.lvmi into tertiles or quartiles
data$log.lvmi_cat <- cut(data$log.lvmi, breaks = quantile(data$log.lvmi, probs = seq(0, 1, by = 0.25)), include.lowest = TRUE)

# Fit the Cox model with stratification by log.lvmi categories
stratified_model <- coxph(Surv(fuyrs, status) ~ sex + age + con.cabg + creat + lv + sten.reg.mix + strata(log.lvmi_cat), data = data, x=T)
summary(stratified_model)

# Check proportional hazards assumption
cox.zph_stratified <- cox.zph(stratified_model)
print(cox.zph_stratified)
```
The new Stratified model now stratifies for the log.lvmi variable, allowing for a more cautious approach towards the Proportional Hazards assumption. Under this model:

1. Sex (Female): The coefficient for females remains non-significant against the base model (p = 0.14608). The hazard ratio is 1.626287, indicating a 62.3% higher risk of death for females compared to males, though this difference is not statistically robust. 2. **Age**: Again, this variable stays significant across the models, each additional year of age increases the risk of death by 10.3% (p = 1.21e-06). Again, this confirms the substantial impact of age on survival post-surgery, reflecting the increased vulnerability and potential comorbidities associated with older age.
3. **Coronary Artery Bypass Grafting (CABG) (Yes)**: The hazard ratio for CABG has slightly is now 2.016968 remaining significant (p = 0.02372).
4. Creatinine (Creat): The effect of creatinine levels remains non-significant (p = 0.13550).
5. **Left Ventricular Function (LV)**: The effect of moderate ventricular function, in this augmented model is now even more significant (p-value=0.03198) and the low ventricular function stays very highly significant (p-value=0.00028) with a very high hazard ratio of 5.422668. Again, low ventricular function is a strong predictor for both models, representing a significant risk factor for mortality post-surgery.
6. **Stenosis Severity (Sten.reg.mix)**: The effect of stenosis severity remains non-significant for regurgitation (p-value=0.12958), while staying statistically significant for mixed conditions (p-value=0.01395). Compared to stenosis, mixed conditions seem to show a protective effect (HR=0.302234, or 30.22% lower risk).

Again, to evaluate the functional form of continuous variables and ensure that the proportional hazards assumption holds for each covariate, we will visually inspect the Schoenfeld Residuals.

Finally, the tests for proportional hazards provided no significant evidence against the assumption for the included variables when stratified by LVMi categories. The global test p-value of 0.42 further supports the appropriateness of the model. For this reason, this model seems to be the most appropriate among the three - this theory will be further investigated while calculating for the main performance indicators.

```{r Schonefeld and Martingale Residuals}
# Evaluate the proportional hazards assumption using the cox.zph function
# This function tests the proportional hazards assumption based on the scaled Schoenfeld residuals
ph_test_2 <- cox.zph(stratified_model)

# Plotting the residuals to visually inspect any trend indicating non-proportionality
plot(ph_test_2)  # This generates multiple plots if there are multiple covariates

# Print the global test of proportional hazards
print(ph_test_2) # Displays test statistics and p-values for each covariate
```

# Performance Evaluation

```{r Base Model predictions}
fit1<-survfit(base_model,newdata=data)
data$risk_base<-1-as.numeric(summary(fit1,times=5)$surv)
```

```{r Augmented Model predictions}
fit2<-survfit(augmented_model,newdata=data)
data$risk_augmented<-1-as.numeric(summary(fit2,times=5)$surv)
```

```{r Augmented Model Stratified predictions}
fit3<-survfit(stratified_model,newdata=data)
data$risk_augmented_strata<-1-as.numeric(summary(fit3,times=5)$surv)
```

### Calibration: Plot and Brier Scores 

```{r Scores for ROC and Calibration Plot}
score <- Score(list("Base"=base_model,"Augmented"=augmented_model,"Stratified"=stratified_model),
              formula=Surv(fuyrs, status==1)~1,
              data=data,conf.int=T,
              times=seq(1,5,1),
              plots=c("calibration","ROC"))

plotCalibration(score,times=5,cens.method="local",method="quantile",q=10)
title(main="Calibration Plot at 5 years")

```
The calibration plot provides a visual assessment of the performance of three Cox proportional hazards models—Base, Augmented, and Stratified—at predicting the 5-year survival outcomes for patients following heart valve implantation.

The three models all indicate a fairly good calibration (with Brier scores between 11.0 and 11.3), even though the Base Model appears to underestimate the event risk more heavily than the other two models, especially at lower risk predictions (particularly at the 25% predicted risk score). This behaviour suggests that additional factors or more complex interactions may need to be considered in order to improve the models' sensitivity in these regions.


### Discrimination: ROC Curves and AUC

```{r Plotting the ROC Curves}
plotROC(score,times=5,cens.method="local")
title(main="time-dependent ROC at 5 years")
```
The Receiver Operating Characteristic (ROC) curve is a critical tool used to evaluate the diagnostic ability of predictive models over a range of thresholds. This time-dependent ROC curve illustrates how well the three Cox proportional hazards models—Base, Augmented, and Stratified—discriminate between patients who experienced an event and those who did not by the 5-year follow-up.

All three models significantly outperform a random classifier, as indicated by the ROC curves residing well above the diagonal line. Again, The curves of the augmented and stratified models are very close, suggesting that both models provide similar discriminative power. The stratified model does not show a clear advantage in discrimination over the augmented model, despite its slightly (apparent) better calibration, that will be investigated more further on.

### Calibration and Discrimination: Brier Scores and AUC (comments on the graphical results)

```{r Scores}
score <- Score(list("Base"=base_model,"Augmented"=augmented_model,"Stratified"=stratified_model),
              formula=Surv(fuyrs, status==1)~1,
              data=data,conf.int=T,
              times=seq(1,5,1))

summary(score)
```

With the Score function it is possible to evaluate and compare the performance of the three Cox proportional hazards models over a 5-year period using the comparison of Brier scores and the Area Under the Curve (AUC) of the Receiver Operating Characteristic (ROC).

The *Brier score* measures the accuracy of probabilistic predictions by evaluating the mean squared difference between predicted probabilities and the actual outcomes. The comparisons between models generally show minor differences, with no significant improvements noted in the augmented or stratified models compared to the base model. This suggests that while stratification and augmentation may affect discrimination (as seen in AUC), they do not significantly enhance the overall accuracy of the probability predictions.

The *AUC Analysis* confirms a general good discriminative ability for all three models (with a notable peak at year 1 for the Stratified model, with an AUC score of 87.3%), although the improvements of both the Augmented and Stratified model over the Base one seem to be statistically non-significant, suggesting modest benefits from including additional variables (log.lvmi) and stratification.

### Internal Validation 

```{r Bootstrapping for internal validation}
Bootscore <- Score(list("Base"=base_model,"Augmented"=augmented_model,"Stratified"=stratified_model),
              formula=Surv(fuyrs, status==1)~1,
              data=data,conf.int=T,
              times=seq(1,5,1),
              split.method="loob",B=100,seed=1000)

summary(Bootscore)
```

In order to secure an internal validation of the model comparison, bootstrap validation has been employed (over the 5-year period) - indicating that a statistical significant improvement in the models (either from a decrease in the Brier Scores or an increase in the AUC) is not present.

In conclusion, by *roughly* accepting the Proportional Hazards assumption as *not violated* for the **Augmented** version of the model, it is possible to interpret it as the most efficient in evaluating and predicting the risks of event.

## Net Benefit

```{r}
library(dcurves)
dca(Surv(fuyrs, status==1) ~ risk_base + risk_augmented + risk_augmented_strata, data, time = 5)
```

The provided graph from the Decision Curve Analysis (DCA) offers a visualization of the net benefit derived from using the base and augmented Cox proportional hazards models to make clinical decisions based on a 5-year survival prediction. DCA is used to evaluate the clinical usefulness of prediction models by comparing the net benefits at various threshold probabilities.

Based on the graph, both the **Base** and **Augmented** Models seem to be potentially beneficial for clinical decision-making within specific ranges of threshold probabilities: in particular, up until a threshold probability of 50% offer improvements in Net Benefit against the "Treat None" scenario.

# Model Prediction: 3 subjects on a 5-year span
```{r}
# Prediction of event risk at a fixed time-point (e.g., 5 years) for 3 "type" subjects
# Create example subjects
type_subjects <- data.frame(
  sex = c("Male", "Female", "Male"),
  age = c(100, 70, 65),
  con.cabg = c("Yes", "No", "Yes"),
  creat = c(100, 150, 120),
  lv = c("High", "Moderate", "Low"),
  sten.reg.mix = c("Stenosis", "Regurgitation", "Mixed"),
  log.lvmi = c(log(150), log(170), log(160))
)

type_subjects_strat <- data.frame(
  sex = c("Male", "Female", "Male"),
  age = c(100, 70, 65),
  con.cabg = c("Yes", "No", "Yes"),
  creat = c(100, 150, 120),
  lv = c("High", "Moderate", "Low"),
  sten.reg.mix = c("Stenosis", "Regurgitation", "Mixed"),
  log.lvmi = c(log(150), log(170), log(160))
)

type_subjects_strat$log.lvmi_cat <- cut(type_subjects_strat$log.lvmi, breaks = quantile(data$log.lvmi, probs = seq(0, 1, by = 0.25)), include.lowest = TRUE)

# Predict event risk at 5 years using both models
risk_basic <- predict(base_model, newdata = type_subjects, type = "risk", se.fit = TRUE)
risk_augmented <- predict(augmented_model, newdata = type_subjects, type = "risk", se.fit = TRUE)
risk_strata <- predict(stratified_model, newdata = type_subjects_strat, type = "risk", se.fit = TRUE)

risk_basic
risk_augmented
risk_strata
```
The three example subjects considered for testing the predictive capabilities of the models are defined as follows:

**Subject 1**: Male, 100 years old, with coronary artery bypass grafting (CABG), creatinine level of 100, high left ventricular function, stenosis, and log(LVMi) of approximately 5.01.

**Subject 2**: Female, 70 years old, without CABG, creatinine level of 150, moderate left ventricular function, regurgitation, and log(LVMi) of approximately 5.14.

**Subject 3**: Male, 65 years old, with CABG, creatinine level of 120, low left ventricular function, mixed valve hemodynamics, and log(LVMi) of approximately 5.08.

The **Base** and **Augmented** models yield similar risk predictions for all subjects, with higher risk observed for the oldest subject (Subject 1), particularly in the augmented model, suggesting a slightly higher sensitivity to the variables included beyond the base model. The risk estimates for Subjects 2 and 3 are relatively low in both models, indicating lesser influence of their characteristics on the outcome according to these models. The **Stratified** model shows notably lower risk predictions for all subjects, especially for Subject 1. This suggests that the stratification based on log(LVMi) categories significantly impacts the model's risk assessment, potentially adjusting for the variable non-proportionality and specific subgroup characteristics.