---
title: "Heart.valve.data"
output:
  html_document:
    df_print: paged
---
# Library

```{r}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(readr)
```
# Importing Data 
```{r}
file_path <- "/Users/alessandrorota/Desktop/LM - DATA SCIENCE/II anno/Data science lab in medicine/Big data in Healthcare/progetto/heart.valve.txt"
heart.valve <- read_delim(file_path, delim = "\t")
head(heart.valve)
```

```{r}
dim(heart.valve)
```
Description of dataset variables:

-   ***Paz.id***: patient identification number.

-   ***log.lvmi***: natural logarithm of the “Left Ventricular Mass Index” (standardized) measured at baseline.

-   ***fuyrs***: follow-up time from surgery, in years.

-   ***status***: event indicator (1 = dead; 0 = censored).

-   ***sex***: sex of the patient (0 = M; 1 = F).

-   ***age***: age of the patient (years) at surgery.

-   ***con.cabg***: presence of coronary bypass (1 = yes; 0 = no).

-   ***creat***: pre-operative serum creatinine (µmol/mL).

-   ***lv***: pre-operative left ventricular ejection fraction (1 = high, 2 = moderate, 3 = low).

-   ***sten.reg.mix***: hemodynamics of the heart valve (1 = stenosis, 2 = regurgitation, 3 =
mixed).


```{r}
str(heart.valve)
```
```{r}
summary(heart.valve)
```

# Age distribution

```{r}
ggplot(heart.valve, aes(x = age)) +
  geom_histogram(binwidth = 5, fill = "blue", color = "black") +
  ggtitle("Distribuzione dell'età") +
  xlab("Età") +
  ylab("Frequenza")
```
#Correlation between variables


```{r}
cor_matrix <- cor(heart.valve, use = "complete.obs")
print(cor_matrix)

if (!require(corrplot)) {
    install.packages("corrplot")
    library(corrplot)
}
corrplot(cor_matrix, method = "circle")
```


## Left Ventricular Mass Index

Il Left Ventricular Mass Index (LVMI) è un parametro clinico utilizzato per valutare la massa del ventricolo sinistro del cuore in relazione alla superficie corporea del paziente. È una misura importante nella diagnosi e nel monitoraggio di condizioni cardiache come l'ipertrofia ventricolare sinistra, che è un ispessimento delle pareti del ventricolo sinistro del cuore.

```{r}
ggplot(heart.valve, aes(x = factor(status), y = log.lvmi, fill = factor(status))) +
  geom_boxplot() +
  ggtitle("Boxplot di log.lvmi per Status") +
  xlab("Status") +
  ylab("log.lvmi") +
  scale_fill_discrete(name = "Status")

```

```{r}
ggplot(heart.valve, aes(x = age, y = log.lvmi)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  ggtitle("Scatter Plot: età vs log.lvmi") +
  xlab("Età") +
  ylab("log.lvmi")

```








